Search for WIMP Inelastic Scattering Off Xenon Nuclei With Xenon100 Data

### 2017-01-16 Inelastic/intro.tex
updated_intro_and_ref

\section{\label{sec:level1} Introduction}
Astrophysical and cosmological evidence indicates that the dominant mass fraction of our Universe consists of some yet unknown
form of dark, or invisible matter. The dark matter could be made of new, yet undiscovered particles, and well-motivated
theoretical models going beyond the Standard Model of particle physics predict the existence of Weakly Interacting Massive
Particles (WIMPs), which are natural candidates for dark matter. This hypothesis is currently being tested by several direct
and indirect detection experiments, as well as at the LHC~\cite{Bertone:2004pz,Baudis:2016qwx}.

@@

We consider the $^{129}\text{Xe}$ isotope, which has an abundance of 26.4\% in natural xenon, and a lowest-lying 
3/2$^{+}$ state at 36.6\,keV above the 1/2$^+$ gound state. The electromnagnetic nuclear decay has a half-life of 0.97\,s. 
The sigantures of inelastic scattering in xenon have been studied in detail in~\cite{Baudis:2013bba}. It was found that this 
channel is complementary to spin-dependent, elastic scattering, dominating the integrated rates above $\simeq10$\,keV energy 
depositions. In addition, in case of a positive signal, the observation of inelastic scattering would provide a clear 
indication of the spin-dependent nature of the fundamental interaction.

\section{The XENON100 Detector}

The XENON100 experiment operated a dual-phase (liquid and gas) xenon time projection chamber (TPC) at the Laboratori Nazionali 
del Gran Sasso (LNGS) in Italy. It contained 161\,kg of xenon in total,  with  62\,kg in the active region of the TPC. These 
were monitored by 242 low-radioactivity, UV-sensitive photomultiplier tubes (PMTs) arranged in two arrays, one in the liquid 
and one in the gas. The PMTs detected the prompt scintillation (S1) and the delayed, proportional scintillation signal (S2) 
created by a particle interacting in the active TPC region. The S2-signal is generated due to ionisation electrons, drifted 
and extracted into the gas phase, where the proportional scintillation, or electroluminiscence, is produced. These photons 
carry the $x-y$ information of the interaction site, while the $z-$information came from the drift time measurement. The TPC 
thus yielded a three-dimensional event localisation, enabling to reject the majority of background events via fiducial volume 
cuts, as well a multiple interactions, which are unlikely due to dark matter particles~\cite{Aprile:2011dd}. The ratio of 
S2/S1 provided the basis for distinguishing between nuclear recoils (NRs), as induced by fast neutrons and expected from 
elastic WIMP-nucleus scatters, and electronic recoils (ERs) produced by $\beta$ and $\gamma$-rays. As we shall see, the 
inelastic scattering signature will be a superposition of an ER and a NR, due to the low lifetime of the excited state and 
the short mean free path ($\sim$0.15\,mm) of the 39.6\,keV photon.

XENON100 has acquired science data between 2008-2015, and has set competitive constraints on spin-
independent~\cite{Aprile:2012nq,Aprile:2016swn} and spin-dependent~\cite{Aprile:2013doa,Aprile:2016swn} elastic WIMP-nucleus 
scatters, as well as on axions/ALPs~\cite{Aprile:2014eoa} and leptophilic dark matter 
models~\cite{Aprile:2015ade,Aprile:2015ibr,Aprile:2017yea}.

>>>>

\section{\label{sec:level1} Introduction}
Astrophysical and cosmological evidence indicates that the dominant mass fraction of our Universe consists of some yet unknown
form of dark, or invisible matter. The dark matter could be made of new, yet undiscovered particles, and well-motivated
theoretical models going beyond the Standard Model of particle physics predict the existence of Weakly Interacting Massive
Particles (WIMPs), which are natural candidates for dark matter. This hypothesis is currently being tested by several direct
and indirect detection experiments, as well as at the LHC~\cite{Bertone:2010zza,Baudis:2016qwx}.

@@

We consider the $^{129}\text{Xe}$ isotope, which has an abundance of 26.4\% in natural xenon, and a lowest-lying 
3/2$^{+}$ state at 36.6\,keV above the 1/2$^+$ gound state. The electromnagnetic nuclear decay has a half-life of 0.97\,s. 
The signatures of inelastic scattering in xenon have been studied in detail in~\cite{Baudis:2013bba}. It was found that this 
channel is complementary to spin-dependent, elastic scattering, dominating the integrated rates above $\simeq10$\,keV energy 
depositions. In addition, in case of a positive signal, the observation of inelastic scattering would provide a clear 
indication of the spin-dependent nature of the fundamental interaction.

Our paper is structures as follows.  In Section~\ref{sec:xenon100} we briefly describe the XENON100 detector and the employed data 
set in this analysis. In Section~\ref{sec:analysis} we detail the data analysis method, including the simulation of the expected signal and the 
background model. We conclude in Section~\ref{sec:results} with our results and new constraints on inelastic WIMP-nucleus scatters.

\section{The XENON100 Detector}
\label{sec:xenon100}

The XENON100 experiment operated a dual-phase (liquid and gas) xenon time projection chamber (TPC) at the Laboratori Nazionali 
del Gran Sasso (LNGS) in Italy. It contained 161\,kg of xenon in total,  with  62\,kg in the active region of the TPC. These 
were monitored by 242 1-inch square, low-radioactivity, UV-sensitive photomultiplier tubes (PMTs) arranged in two arrays, one in the liquid 
and one in the gas. The PMTs detected the prompt scintillation (S1) and the delayed, proportional scintillation signal (S2) 
created by a particle interacting in the active TPC region. The S2-signal was generated due to ionisation electrons, drifted 
in a field of 530\,V/cm and extracted into the gas phase by a field of 12\,kV/cm, where the proportional scintillation, or electroluminiscence, 
was produced. These photons carried the $x-y$ information of the interaction site, while the $z-$information came from the drift time measurement. The TPC 
thus yielded a three-dimensional event localisation, enabling to reject the majority of background events via fiducial volume 
cuts, as well a multiple interactions, which are unlikely due to dark matter particles~\cite{Aprile:2011dd}. The ratio of 
S2/S1 provided the basis for distinguishing between nuclear recoils (NRs), as induced by fast neutrons and expected from 
elastic WIMP-nucleus scatters, and electronic recoils (ERs) produced by $\beta$ and $\gamma$-rays. As we shall see, the 
inelastic scattering signature will be a superposition of an ER and a NR, due to the low lifetime of the excited state and 
the short mean free path ($\sim$0.15\,mm) of the 39.6\,keV photon.

XENON100 has acquired science data between 2008-2015, and has set competitive constraints on spin-independent~\cite{Aprile:2012nq,Aprile:2016swn} 
and spin-dependent~\cite{Aprile:2013doa,Aprile:2016swn} elastic WIMP-nucleus 
scatters, as well as on axions/ALPs~\cite{Aprile:2014eoa} and leptophilic dark matter models~\cite{Aprile:2015ade,Aprile:2015ibr,Aprile:2017yea}.

Here

### 2017-01-16 Inelastic/results.tex
updated_intro_and_ref

\section{Results}
Using an exposure of 34~kg of liquid xenon and 224.6~live days of data a yield of 764 events is observed  in the region of interest,
this is compatible with the expectation of $756 \, \pm \, 5^{(stat.)} \, \pm 55^{(syst.)}$ events from the background only hypothesis. 
Figure~\ref{fig:dataVSbkg} shows how these events are distributed in the region of interest, the bottom panel shows the ratio
between data and expected background, where the gray and orange shaded areas represent respectively statistical and systematic uncertainty 
on the background expectation.
%The distribution of events from data are compared to the background expectation in Figure~\ref{fig:dataVSbkg} and with the relative
%uncertainties.

>>>>

\section{Results}
\label{sec:results}

Using an exposure of 34~kg of liquid xenon and 224.6~live days of data a yield of 764 events is observed  in the region of interest,
this is compatible with the expectation of $756 \, \pm \, 5^{(stat.)} \, \pm 55^{(syst.)}$ events from the background only hypothesis. 
Figure~\ref{fig:dataVSbkg} shows how these events are distributed in the region of interest, the bottom panel shows the ratio
between data and expected background, where the gray and orange shaded areas represent respectively statistical and systematic uncertainty 
on the background expectation.
%The distribution of events from data are compared to the background expectation in Figure~\ref{fig:dataVSbkg} and with the relative
%uncertainties.
